---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: # Rep_Res_PA1.html
    keep_md: TRUE
---


## Loading and preprocessing the data
```{r load and preprocess }
library(knitr)
cache=TRUE
act <- read.csv(file="activity.csv") # Read CSV (steps, date, interval)
udates<-unique(act[,2])  # Get Unique dates
tsteps<-c()  # Init total steps, hist steps & count of NA
hsteps<-c()
nac<-c()
for (i in 1:61) {
  ssteps<-act[,1][act[,2]==udates[i]]   # extract single date steps
  nac<-rbind(nac,sum(is.na(ssteps)))
  hsteps <- c(hsteps,sum(ssteps)) # histo steps for each date
  dsteps <- c(udates[i],sum(ssteps),mean(ssteps), median(ssteps))
  tsteps <- rbind(tsteps,dsteps) # total steps data frame
}
```

## What is mean total number of steps taken per day?
### 1. Histogram of total #of steps per day

```{r Histogram of Sum Steps}
hist (hsteps, main="Histogram - Sum Steps")
```

### 2. Mean and Median of total # of steps per day
```{r mean and median of steps, results="asis"}
library(xtable)
dtab<-xtable(summary(tsteps[,3:4]))# get to table format
print (dtab, type="html")
```

## What is the average daily activity pattern?

### 1. Time series plot of 5 min interval data

```{r Time Series 5 min }
plot(act[,1] ~ act[,3], type="l", main="Time Series (5 min)", xlab="Interval",
ylab="steps")
```

### 2. Time Series plot of daily average data
```{r Time Series Daily Avg data}
plot(tsteps[,3] ~ tsteps[,1], type="l", main="Time Series (Avg)", xlab="Interval",
ylab="steps")
```


## Imputing missing values
### 1. Total #of Missing values (Total #of NAs in rows)
```{r #of NAs in rows }
totNA <- sum(rowSums(is.na(act))) # # of NAs in rows
```

Number of NAs in rows is `r totNA`.

### 2. Imput Missing NAs with mean Value
### 3. New Dataset with NAs removed
```{r replace NA by mean of column }
actna<-act     # copy act to actna (NA removed)
for (i in 1:61) {
  ssteps<-actna[,1][actna[,2]==udates[i]]   # extract single date steps
  if (sum(is.na(ssteps) != 0)) {
     actna[,1][is.na(actna[,1])] <- mean(actna[,1], na.rm = TRUE) # to mean
     ssteps<-actna[,1][actna[,2]==udates[i]]    } # re calculate single date steps  
  hsteps <- c(hsteps,sum(ssteps)) # histo steps for each date
  dsteps <- c(udates[i],sum(ssteps),mean(ssteps), median(ssteps))
  tsteps <- rbind(tsteps,dsteps) # total steps data frame
}
```
### 4. Histogram of total #of steps per day, 
### Mean and median of new data set (NAs removed)

``` {r Histogram of Sum Steps - NAs removed}
hist (hsteps, main="Histogram - Sum Steps - NAs removed")
```
  

### Are there differences in activity patterns between weekdays and weekends?

``` {r mean and median compare }
dsteps1<-c()
tsteps1<-c()
for (i in 1:61) {
  ssteps1<-act[,1][act[,2]==udates[i]]   # extract single date steps
  ssteps2<-actna[,1][actna[,2]==udates[i]]   # extract single date steps
  dsteps1 <- c(udates[i],mean(ssteps1), median(ssteps1),mean(ssteps2), median(ssteps2))
  tsteps1 <- rbind(tsteps1,dsteps1) # total steps data frame
}
```

### Weekdays and Weekend activities
``` {r Wkdays and Wkends compare }
actna$day <- format(as.Date(actna[,2]),"%w") # get day number
actna$wkd <- ifelse(( actna$day==0) | (actna$day ==6), "WKEND","WKDAY")
```
### . Time Series plot of daily average data
```{r Time Series Daily Avg WkDay/WkEnd}
library(lattice)
xyplot(actna[,1] ~ actna[,3] | actna[,5], type="l",xlab="Interval",ylab="Steps")

```

